IMPC Mouse data - Heterogeneity

### 2019-07-25 R/Heterogeneity.Rmd
Add files via upload. Need to reconcile with Heterogeneity script, which will probably now become part of the main ms, not supplement

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne & Felix Zajitschek"
date: "2nd April 2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
runtime: shiny
---

@@

```{r}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
```

# Loading functions that are necessary
### Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a cleaned up file which is less computing intensive. However, cvs
```{r}

@@

  mydata[i,] %>% 
  select(strain_name, production_center, trait) %>%
    mutate(
      effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )

@@

 Function to calculate meta-analysis statistics. Created by A M Senior @ the University of Otago NZ 03/01/2014

Below are functions for calculating effect sizes for meta-analysis of variance. 
All functions take the mean, sd and n from the control and experimental groups.

The first function, Cal.lnCVR, calculates the the log response-ratio of the coefficient of variance (lnCVR) - see Nakagawa et al 2015.

The second function calculates the measurement error variance for lnCVR. As well as the aforementioned parameters, this function also takes
Equal.E.C.Corr (default = T), which must be True or False. If true, the function assumes that the correlation between mean and sd (Taylor's Law) 
is equal for the mean and control groups, and, thus these data are pooled. If False the mean-SD correlation for the experimental and control groups
are calculated separately from one another.

@@

Calc.var.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN, Equal.E.C.Corr=T) {
  if(Equal.E.C.Corr==T){
    mvcorr <- 0 #cor.test(log(c(CMean, EMean)), log(c(CSD, ESD)))$estimate   old, slightly incorrect
    S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * mvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * mvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))
  }
  else{
    Cmvcorr <- cor.test(log(CMean), log(CSD))$estimate
    Emvcorr <- cor.test(log(EMean), (ESD))$estimate
    S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * Cmvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * Emvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))		
  }
  S2
}

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than the .csv - which we can still provide for those who absolutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")

data1 <- readRDS("export/data_clean.rds")
#data1 
```

### Clean data
This requires the selection of traits that have been measured in at least 2 centers. Consequently, rare or unusual methods and procedures are being filtered out in this step.
```{r }
dat1 <-
  data1 %>%
  group_by(parameter_name) %>%
  summarize(center_per_trait = length(unique(production_center, na.rm = TRUE)))

@@

data2 <- dat_moreThan1center
#min(data2$center_per_trait)  # as a check if there indeed are no single occurences

@@

## Preparation for Meta analysis
### Define population variable & add grouping variables

@@

 group <- read.csv("export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

@@

### Assign each unique parameter_name (=trait,use trait variable) a unique number ('id')
We add a new variable, where redundant traits are combined
[note however, at this stage the dataset still contains nonsensical traits, i.e. traits that may not contain any information on variance]

```{r}
#head(data)

@@

### Create a matrix to store results for all traits
As the current version of this script utilizes a loop instead of tidyR code, it is here necessary to create an empty matrix, in which the returning values will be stored.

```{r}
results.alltraits.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*14), ncol = 14))) #number of individual results per trait = 10
names(results.alltraits.grouping) <- c("id", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se" , "sampleSize", "trait")
```

## LOOP, to run meta-analysis on all traits
The loop combines the functions mentioned above and fills the data matrix with results from our meta analysis. 
Error messages indicate traits that either did not reach convergence, or that did not return meaningful results in the meta-analysis, due to absence of variance. Those traits will be removed in later steps, outlined below.

#lnCVR,  log repsonse-ratio of the coefficient of variance    
    cvr <- metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(~1| strain_name, ~1|production_center, ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000),  verbose=F, data = results)

    results.alltraits.grouping[t, 2] <- cvr$b
    results.alltraits.grouping[t, 3] <- cvr$ci.lb
    results.alltraits.grouping[t, 4] <- cvr$ci.ub
    results.alltraits.grouping[t, 5] <- cvr$se

    cvr

    #lnVR, comparison of standard deviations

cv <- metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(~1| strain_name, ~1|production_center,~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000),  verbose=F, data = results)

    results.alltraits.grouping[t, 6] <- cv$b
    results.alltraits.grouping[t, 7] <- cv$ci.lb
    results.alltraits.grouping[t, 8] <- cv$ci.ub
    results.alltraits.grouping[t, 9] <- cv$se

    # for means, lnRR

means <- metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR, random = list(~1| strain_name, ~1|production_center, ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), data = results)
    results.alltraits.grouping[t, 10] <- means$b
    results.alltraits.grouping[t, 11] <- means$ci.lb
    results.alltraits.grouping[t, 12] <- means$ci.ub
    results.alltraits.grouping[t, 13] <- means$se

        results.alltraits.grouping[t, 14] <- means$k
        results.alltraits.grouping[t, 15] <- unique(results$trait)

# Now that we have a "results" table with each of the meta-analytic means for all effect sizes of interest, we can use this table as part of the Shint App, which will then be able to back calculate the percentage differences between males and females for mean, variance and coefficient of variance. We'll export and use this in the Shiny App. **Note that I have not dealt with convergence issues in some of these models, and so, this will need to be done down the road**

trait_meta_results <- write.csv(results.alltraits.grouping, file = "export/trait_meta_results.csv")

@@

### Merging datasets

Procedure names, grouping variables etc. are merged back together with the results from the metafor analysis above.
This requires loading of another excel sheet, "procedures.csv"

```{r}
procedures <- read.csv("export/procedures.csv")

results.alltraits.grouping$parameter_group <- data$parameter_group[match(results.alltraits.grouping$id, data$id)]
results.alltraits.grouping$procedure <- data$procedure_name[match(results.alltraits.grouping$id, data$id)]

results.alltraits.grouping$GroupingTerm <-  procedures$GroupingTerm[match(results.alltraits.grouping$procedure, procedures$procedure)]
results.alltraits.grouping$parameter_name <- data$parameter_name[match(results.alltraits.grouping$id, data$id)]

meta1 <- results.alltraits.grouping
n <- length(unique(meta1$parameter_name)) # 232
```

Removal of traits that did not achieve convergence, are nonsensical for analysis of variance (such as traits that show variation, such as number of ribs, digits, etc). 14 traits from the originally 232 that had been included are removed.

```{r}
meta_clean <- meta1[ !(meta1$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
removed <-length(unique(meta_clean$parameter_name)) #218
```

## Dealing with Correlated Parameters

This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example, hierarchical parametrization within immunological assays). As those data-points are not independent of each other,  we conducted a meta analyses on these correlated parameters to collapse the number of levels.

### Collapsing correlated parameters

Here we double check numbers of trait parameters in the dataset

meta1 <- meta_clean 
# length(unique(meta1$procedure)) #18
# length(unique(meta1$GroupingTerm)) #9 
# length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name" 
length(unique(meta1$parameter_name)) #218

### Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size) 
This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same  parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

kable(cbind (meta1 %>%  count(parameter_group) )) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

meta1b <-
  meta1 %>%
  group_by(parameter_group) %>% 
  summarize(par_group_size = length(unique(parameter_name, na.rm = TRUE)))
#this gives a summary of number of parameter names in each parameter group, now it neeeds to get merged it back together

meta1$par_group_size <- meta1b$par_group_size[match(meta1$parameter_group, meta1b$parameter_group)]

# Create subsets with > 1 count (par_group_size > 1)

meta1_sub<-subset(meta1,par_group_size >1) # 90 observations   
meta1_sub$sampleSize <- as.numeric(meta1_sub$sampleSize)

# Preparation for Visualization and Meta-analysis

### Perform meta-analyses on correlated sub-traits, using `robumeta` 
The subset of the data is prepared (nested), and in this first step the model of the meta analysis effect sizes are calculated

```{r}
# nesting
n_count <- meta1_sub %>% 
  group_by(parameter_group) %>% 
  mutate(raw_N = sum(sampleSize)) %>%  
    nest()

model_count <- n_count %>% 
  mutate(model_lnRR = map(data, ~ robu(.x$lnRR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnRR_se)^2 )),
  model_lnVR = map(data, ~ robu(.x$lnVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnVR_se)^2 )),
  model_lnCVR = map(data, ~ robu(.x$lnCVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnCVR_se)^2 ))) 
```

### Extract and save parameter estimates 
Function to collect the outcomes of the "mini" meta analysis
```{r}
count_fun <- function(mod_sub)
  return(c(mod_sub$reg_table$b.r, mod_sub$reg_table$CI.L, mod_sub$reg_table$CI.U, mod_sub$reg_table$SE))   #estimate, lower ci, upper ci, SE
```

Extraction of values created during Meta analysis using robu meta
```{r}
robusub_RR <- model_count %>% 
  transmute(parameter_group, estimatelnRR = map(model_lnRR, count_fun)) %>% 
  mutate(r = map(estimatelnRR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnRR) %>%
  purrr::set_names(c("parameter_group","lnRR","lnRR_lower","lnRR_upper","lnRR_se"))

robusub_CVR <- model_count %>% 
  transmute(parameter_group, estimatelnCVR = map(model_lnCVR, count_fun)) %>% 
  mutate(r = map(estimatelnCVR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnCVR) %>%
  purrr::set_names(c("parameter_group","lnCVR","lnCVR_lower","lnCVR_upper","lnCVR_se"))

robusub_VR <- model_count %>% 
  transmute(parameter_group, estimatelnVR = map(model_lnVR, count_fun)) %>% 
  mutate(r = map(estimatelnVR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnVR) %>%
  purrr::set_names(c("parameter_group","lnVR","lnVR_lower","lnVR_upper","lnVR_se"))

robu_all <- full_join(robusub_CVR, robusub_VR) %>% full_join(., robusub_RR)
kable(cbind(robu_all, robu_all)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

 Merge the two data sets (the new [robu_all] and the initial [uncorrelated sub-traits with count = 1]) 
```{r}
meta_all <- meta1 %>% filter(par_group_size == 1) %>% as_tibble
#str(meta_all)
#str(robu_all)
#which(is.na(match(names(meta_all),names(robu_all))))  # check
```

Combine data 
```{r}
# Step1 
combinedmeta <- bind_rows(robu_all, meta_all)
#glimpse(combinedmeta)

# Steps 2&3
metacombo <- combinedmeta
metacombo$counts <- meta1$par_group_size[match( metacombo$parameter_group, meta1$parameter_group)]
metacombo$procedure2 <-meta1$procedure[match( metacombo$parameter_group, meta1$parameter_group)]
metacombo$GroupingTerm2 <-meta1$GroupingTerm[match( metacombo$parameter_group, meta1$parameter_group)]

kable(cbind (metacombo, metacombo)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

```
### Clean-up and rename 
```{r}
metacombo <-metacombo[, c(1, 20:22, 2:13)] 
names(metacombo)[3] <- "procedure" 
names(metacombo)[4] <- "GroupingTerm"

# Quick pre-check before doing plots

compare <- metacombo %>%
  group_by(GroupingTerm) %>%
  dplyr::summarize(MeanCVR = mean(lnCVR),MeanVR = mean(lnVR), MeanRR = mean(lnRR) )

compare

## Last step: meta-meta-analysis 
### Perform meta-analyses (3 for each of the 9 grouping terms: lnCVR, lnVR, lnRR) 
This is the full result dataset
```{r}

kable(cbind (metacombo,metacombo)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

### Prepare data
Nesting, calculating the number of parameters within each grouping term, and running the meta-analysis
```{r}
metacombo_final <- metacombo %>%
 group_by(GroupingTerm) %>%
 nest()

# **Calculate number of parameters per grouping term

metacombo_final <- metacombo_final  %>%  mutate(para_per_GroupingTerm = map_dbl(data, nrow))

# For all grouping terms
metacombo_final_all <- metacombo %>%
 nest()

overall1 <- metacombo_final %>%

  mutate(model_lnCVR = map(data, ~ metafor::rma.uni(yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)),
	   model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)),
	   model_lnRR = map(data, ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)))

# **Final fixed effects meta-analyses ACROSS grouping terms, with SE of the estimate

overall_all1 <- metacombo_final_all %>%

  mutate(model_lnCVR = map(data, ~ metafor::rma.uni(yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)),
	   model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)),
	   model_lnRR = map(data, ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)))

Re-structure data for each grouping term; delete unused variables

```{r}
Behaviour <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Behaviour")    %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Immunology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Immunology")    %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Hematology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Hematology")    %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Hearing <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Hearing")   %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Physiology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Physiology")  %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Metabolism <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Metabolism")  %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Morphology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Morphology")  %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Heart <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Heart")  %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

Eye <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Eye")  %>%  mutate(lnCVR=.[[4]][[1]]$b, lnCVR_lower=.[[4]][[1]]$ci.lb, lnCVR_upper=.[[4]][[1]]$ci.ub, lnCVR_se=.[[4]][[1]]$se,
	              lnVR=.[[5]][[1]]$b, lnVR_lower=.[[5]][[1]]$ci.lb, lnVR_upper=.[[5]][[1]]$ci.ub, lnVR_se=.[[5]][[1]]$se,
		        lnRR=.[[6]][[1]]$b, lnRR_lower=.[[6]][[1]]$ci.lb, lnRR_upper=.[[6]][[1]]$ci.ub, lnRR_se=.[[6]][[1]]$se) )[, c(1,7:18)]

All <- as.data.frame(overall_all1  %>%  mutate(lnCVR=.[[2]][[1]]$b, lnCVR_lower=.[[2]][[1]]$ci.lb, lnCVR_upper=.[[2]][[1]]$ci.ub, lnCVR_se=.[[2]][[1]]$se,lnVR=.[[3]][[1]]$b, lnVR_lower=.[[3]][[1]]$ci.lb, lnVR_upper=.[[3]][[1]]$ci.ub, lnVR_se=.[[3]][[1]]$se,
		        lnRR=.[[4]][[1]]$b, lnRR_lower=.[[4]][[1]]$ci.lb, lnRR_upper=.[[4]][[1]]$ci.ub, lnRR_se=.[[4]][[1]]$se) )[, c(5:16)]

All$lnCVR <- as.numeric(All$lnCVR)
All$lnVR <- as.numeric(All$lnVR)
All$lnRR <- as.numeric(All$lnRR)
All <- All %>% mutate(GroupingTerm = "All")

overall2 <- bind_rows(Behaviour, Morphology, Metabolism, Physiology, Immunology, Hematology, Heart, Hearing, Eye, All)

# PLOTS

## Plot FIGURE 2 (First-order meta analysis results)

Re-order grouping terms

meta_clean$GroupingTerm <- factor(meta_clean$GroupingTerm, levels =c("Behaviour","Morphology","Metabolism","Physiology","Immunology","Hematology","Heart","Hearing","Eye") )
meta_clean$GroupingTerm <- factor(meta_clean$GroupingTerm, rev(levels(meta_clean$GroupingTerm)))

# *Prepare data for all traits

meta.plot2.all <- meta_clean %>% select(lnCVR, lnVR, lnRR, GroupingTerm) %>% arrange(GroupingTerm)

meta.plot2.all.b <- gather(meta.plot2.all, trait, value, c(lnCVR, lnVR, lnRR))

meta.plot2.all.b$trait <- factor(meta.plot2.all.b$trait, levels =c("lnCVR","lnVR","lnRR") )

meta.plot2.all.c <- meta.plot2.all.b  %>%
				group_by_at(vars(trait, GroupingTerm)) %>%
				summarise(malebias = sum(value > 0), femalebias = sum(value<= 0), total= malebias + femalebias, 
					malepercent = malebias*100/total, femalepercent = femalebias*100/total)

meta.plot2.all.c$label <- "All traits"

# restructure to create stacked bar plots

meta.plot2.all.d <- as.data.frame(meta.plot2.all.c)
meta.plot2.all.e <- gather(meta.plot2.all.d, key = sex, value = percent, malepercent:femalepercent, factor_key = TRUE)

# create new sample size variable

meta.plot2.all.e$samplesize <- with(meta.plot2.all.e, ifelse(sex == "malepercent", malebias, femalebias) )

>>>>

---
title: "IMPC Mouse data - Heterogeneity"
author: "Susanne & Felix Zajitschek"
date: "4rth April 2019"
output:
  html_document:
   code_folding: hide
   toc: true
   toc_depth: 2
---

@@

```{r}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
```
####################################################################################################################################

# ** FIGURE 4 (Second-order meta analysis on heterogeneity)

# Loading functions that are necessary
##Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a the cleaned up file which is less computing intensive. However, cvs
```{r}

@@

  mydata[i,] %>% 
  select(strain_name, production_center) %>%
    mutate(
      effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )

@@

 Function to calculate metanalaysis statistics. Created by A M Senior @ the University of Otago NZ 03/01/2014

Below are funcitons for calculating effect sizes for meta-analysis of variance. 
All functions take the mean, sd and n from the control and experimental groups.

The first function, Cal.lnCVR, calculates the the log repsonse-ratio of the coefficient of variance (lnCVR) - see Nakagawa et al 2015.

The second function calculates the measuremnt error variance for lnCVR. As well as the aforementioned parameters, this function also takes
Equal.E.C.Corr (default = T), which must be True or False. If true, the funciton assumes that the correlaiton between mean and sd (Taylor's Law) 
is equal for the mean and control groups, and, thus these data are pooled. If False the mean-SD correlation for the experimental and control groups
are calculated seperatley from one another.

@@

Calc.var.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN, Equal.E.C.Corr=T) {
  if(Equal.E.C.Corr==T){
    mvcorr <- cor.test(log(c(CMean, EMean)), log(c(CSD, ESD)))$estimate
    S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * mvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * mvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))
  }
  else{
    Cmvcorr <- cor.test(log(CMean), log(CSD))$estimate
    Emvcorr <- cor.test(log(EMean), (ESD))$estimate
    S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * Cmvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * Emvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))		
  }
  S2
}

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can ve loaded and cleaned using the da cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than th e .csv - which we can still provide for those who absoliutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")

data1 <- readRDS("export/data_clean.rds")

## Clean data
This requires the selection of traits that have been measured in at least 2 centers. Consequently, rare or unusual methods and procedures are being filtered out in this step.
```{r }
dat1 <-
  data1 %>%
  group_by(parameter_name) %>%
  summarize(center_per_trait = length(unique(production_center, na.rm = TRUE)))

@@

data2 <- dat_moreThan1center
#min(data2$center_per_trait)  # as a check ifthere indeed are no single occurences

@@

#Preparation for Meta analysis
## Define population variable & add grouping variables

@@

 group <- read.csv("ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

@@

## Assign each unique parameter_name (=trait,use trait variable) a unique number ('id')
We add a new variable, where redundant traits are combined
[note however, at this stage the dataset still contains non-sensical traits, i.e. traits that may not contain any u=information on variance]

```{r}
head(data)

@@

## *Create matrix to store results for all traits
```{r}
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*30), ncol = 30))) 
names(results.allhetero.grouping) <- c("id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR", 
	"s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR", 
	"s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR", 
	"s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", 
	"lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se")

@@

# **LOOP
#  parameters to extract from metafor (sigma2's, s.nlevels)
